How to Recognize a Bad ``Bad Smell'' (using Contrast Sets)

### 2016-03-08 Manuscript.tex
Updates from ShareLaTeX

\usepackage{stfloats}

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{ 0eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\title{Learning Which Bad Smells are 
Real Stinkers?}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract}

\end{abstract}

\keywords{Defect prediction; Configuration; Planning; Decision trees.}
\section{Introduction}

@@

\input{tex/jur_char.tex}

The thesis of this paper is that the XTREE tool can stop developers
wasting time on useless software refactorings. Before presenting experiments
that supports that claim, this section motivates the need for 
such tools.

In many fields,  practitioners hang on to persistent beliefs, even
if the evidence for those beliefs is not strong. For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
There is growing evidence that amongst programmers,
many of their  widely-held beliefs are inaccurate and misleading. 
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer two
specific examples
were  widely-held views in SE that are   now questionable, given recent results. 
Firstly, Nagappan et al. analyzed~\cite{mei15} 
how the {\tt goto} statement was used  in 11,000 Github repositories, 
they found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.

Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. Boehm and Basili comment that this phase delay effect
``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
After reading the literature, and surveying academics
and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
this effect seems now heavily mitigated by $21^{st}$ software development languages,
tools and development practices.  When we mined the 171 software projects shepherded
by the SEI in the period 2006-2014, we found that phase delay results
in only minor changes in the 
time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.

These two studies are symptomatic
of a widespread malaise in software engineering.
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  `` (a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

One area where poorly-justifed beliefs can waste valuable project resources is
refactoring in order to remove code bad smells.

>>>>

%\usepackage{stfloats}

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{ amssymb }
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{ 0eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

@@

\title{  How to Recognize  a Bad  ``Bad Smell'' (using Contrast Sets)}
\numberofauthors{2} 
\author{  
\alignauthor
Rahul~Krishna, Tim Menzies  \\
       \affaddr{Computer Science, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies\}@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}}
\maketitle
\begin{abstract} 
It is hard to assess
if a bad smell is ``bad''; i.e. if it is irrelevant or misleading
for a specific project.
Developers, text books, and tools
disagree on which bad smells are important.
Specifically, it is unclear at what threshold values
for which metrics should trigger code reorganization since
most studies do not offer
metrics {\em and} a threshold for their proposed bad smells.
This paper proposes the XTREE contrast set
learner to assess bad smells
and reject any that are not supported by
evidence in the historical record.  
Using XTREE, we   show that the range of useful
``bad smells'' changes from project to project. Hence,
before refactoring, it is important to run a critique tool
like XTREE to stop work on useless code reorganizations.
Further, using XTREE, it is possible to show that many
of the bad smells endorsed by standard texts or tools
are not useful for particular projects. 
Based on these results, we endorse  
guiding code reorganization via bad smells. However,  
it is important to XTREE the bad smells defined for a 
particular project.

{\bf Categories/Subject Descriptors:} D.2 [Software Engineering]; I.2 [Artificial Intelligence];

{\bf Keywords:} Bad smells,
performance prediction,  decision trees 
\end{abstract}

\section{Introduction}

% Developers, tools, and research papers are either vague or disagree
% about what code refactorings are most important to perform, or ignore.
% As noted by Shantnawi~\cite{Shatnawi10} there is a dearth of
% studies that formulate 
% precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
% claim that OO design metrics (e.g. depth of inheritance,
% classes per method, etc) can be used to predict
% code quality. Yet they their paper never comments at what point
% a particular code measurement turns from ``ok'' to ``bad smell''.
% Similary, code smell detection tools work for a certain
% range of

% When there are many possible code refactorings, it becomes difficult
% to choose which ones to execute. 
% Precious project resources can
% be wasted if developers spend too much time on useless refactorings.

\noindent
According to   Fowler~\cite{fowler99}, bad smells (a.k.a. code smells)
are ``a surface indication that usually corresponds to a deeper problem''.
Fowler strongly recommends   removing   code smells   by
\begin{quote}
``...applying a series of small behavior-preserving transformations, each 
of which seem ``too small to be worth doing''. 
The  effect of   these transformations is quite significant. By doing them in small steps you {\em reduce the risk of introducing errors}'' (our italics).
\end{quote}
The literature is divided on the merits of bad smells.
Much research  endorses using them as a guide
code change. For example,
the recent literature review of Tufano et al. at ASE'15~\cite{Tufano2015}  
lists dozens of  
papers proposing bad smell detection and repair tools.

On the other hand, other papers cast doubts on the value of use bad smalls
as triggers for code change.
In our own work, documented later in this paper, we  show
that tools and text books and developers can disagree on what bad smells
are important enough to hunt down and eliminate. Our findings
are consistent with several other papers.
Mantyla et al.~\cite{Mantyla2004} found that within an organization,
 developers had different perceptions on the importance of different
 bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
 surveyed 85 professional developers to conclude that it remains
 open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
 Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
 performing maintenance tasks that took three to four weeks to complete. 
 After modifying nearly 300 Java files, it was observed
 that none of the dozen investigated code smells explored in 
 that study were mostly irrelevant to to that maintenance work.

Debates on when/what code should be restructed are interesting, especially
in the current context of  widely-used SE practices such as
test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
In that approach, developers are always asking themselves ``should I reorganize this code''?
Bad smells are a widely used  technique for deciding when to reorganize\footnote{Fowler textbook is widely read-- as of  early 2016, ACM Digital Portal reports 
it has 6434 references. Also, bad smell-based research
is widely reported-- a search  on {\em "bad smell"+programming} 
where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
Hence it is important to understand the value of
 bad smells.  
Valuable development resources can be wasted if developers
change their code based on   bad smells that are ``bad''; i.e. 
 irrelevant or misleading
for some particular project.

It turns out that it is hard to learn ``good'' bad smell detectors. Such a detector
is a combination of some code metric (e.g. lines of code in a method)
and a threshold measurement (e.g. {\em loc} $>$ 100). 
Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
discuss what OO design metrics might defect bad smells (e.g. a excessive
depth of inheritance tree) but they never comment on what thresholds
means that some section of code starts to smell bad. After
an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
of  studies  conducted to formulate bad smell guidelines,
represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.

Why are there so few papers offer bad smell metrics and thresholds? For the few that
do exist, why do developers disagree on the relevance of those bad smells?
As discussed on the next page, one reason for this is that 
   cognitive biases result in different
developers declaring that
bad smell X or Y is most important.

When developers cannot fully assess the value of bad smells,
it is possible to use automatic means. We assume that
developers will change code until a bad smell detector stops triggering.
That is, if the bad smell detector is {\em loc} $>$ 100, they will
redistribute the functionality of a long method into several smaller ones.
Recalling the italics in the above quote, Fowler's criteria for successful bad smell removal is the extent
to which the reorganized code has fewer errors/defects.
Data miners can be used to predict the defects in the reorganized code
so, in this paper, we will use static code defect predictors to assess
bad smells.

After making that assessment, this paper finds
that   state-of-the-art methods
(Shantnawi~\cite{Shatnawi10}  in TSE 2010;
and our own instance-based systems~\cite{me12c} from 2012) 
are less effective for bad smell assessment than a new
contrast set learning method called XTREE.  
Standard learners, such as the regression methods used by Shantnawi~\cite{Shatnawi10}
only comment on current distributions.  
  Contrast set learners   recommend {\em what to change} in order to move from
     one classification to another (e.g. from defective=yes to
     defective=no).
     When applied to defect predictors from static code measures,
that delta is a change to static code measures;
i.e. a code reorganization. Hence, constrast set learners
are a better way to learn when code needs to be reorganized
than, say, the methods of Shantnawi~\cite{Shatnawi10}.

We recommend using XTREE as a judge to rank the value of a set
of bad smell detectors:
\begin{itemize}
    \item Consider $N$ developers each proposing a software reorganization
based on a set of bad smells $S$.
\item XTREE will (a)~learn contrast sets
from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
 on any member of $S_{\mathit{bad}} \in S$
that XTREE could {\em not} be found in the historical record.
\end{itemize}
Note that we    recommend  XTREE  is used to critize, but  not to create,
bad smells. 
XTREE uses frequency counts of code features seen in the historical record
to generate
predictions about the effects of changing code
attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
where spurious correlations can suggest effects that are not
truly causal.
That said, if someone else's proposed bad smell detector $S_{\mathit{bad}} $ falls {\em outside} of the space of
effects found  using XTREE's analysis, then we can say that XTREE found 
{\em neither} correlation {\em or}
causal evidence that $S_{\mathit{bad}} $ can be found in the historical record.

XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
by developers. XTREE does not perform those refactorings.
Secondly, XTREE assumes   Fowler's bad smells
should be assessed via their contribution to software defects.   Some researchers 
 comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
about interactions within the code. In that view, 
the concept of refactoring can be useful
even if it does not address issues of defects. 
That said, using
defect predictors,
it is possible to operationalize XTREE
while it might be harder to otherwise operationalize the assessment of bad smells.
Hence, 
this paper assumes that {\em better} bad smells and
bettering at finding code changes that  {\em reducing} the chances
that the new code will contain defects.

The contributions of this paper are
\begin{itemize}

    \item A new kind of contrast set learner called XTREE
    that can critique plans of what to change within a data set in order
    to improve outcomes.
    \item An experimental rig for assessing the value of tools like
    XTREE that propose changes to software. That rig uses data mining
    to build an oracle for the proposed changes.
    \item Using that learner and that oracle, we offer a method for avoid code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
    \item A resolution between proponents of bad smells (e.g. Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}). We say
    that bad smells can guide code change, but only if those bad smells are
    first carefully assessed. 
\end{itemize}

The rest of this paper discusses the cognitive
biases which make developers disagree on the relevance of those bad smells to
their particular projects.  
 That discussion is our motivation for developing the XTREE  tool
that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.

@@

Why build tools like XTREE to  critique proposed developer actions?
Are not those developers the best guide to what is important, or irrelevant
to their code?

It turns our that developer {\em cognitive biases} can mislead them into
asserting that some things are important and relevant when they are not.
This section discusses those biases-- firstly in the general sense
across all SE beliefs; then secondly in the   context of code smells.

A widespread malaise in software engineering is that
beliefs about software are rarely revised and hence may be
  inaccurate and 
misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
strong views on many issues, including bad smells, but those views may be wrong for the current
project.
It should be noted that  software engineering is not the only field with this problem.
  For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10). Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}. Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.

As in medicine, so too in software engineering.
According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects. They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.  J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning. They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.

To support the last paragraph, we offer several
specific studies
showing that   widely-held views in SE that are   now questionable, given recent results. 
The first few come from the general SE literature while the rest
relate specifically to bad smells:
\input{tex/refactor}
\begin{enumerate} 
\item
Nagappan et al. analyzed~\cite{mei15} 
how   {\tt goto}  was used  in 11,000 Github repositories. 
They found that developers
rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
\item
In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
the ``phase delay'' truism that {\em fixing a requirements error  during
coding can be thousands of times more expensive than recognizing and fixing that error
as soon as it arises}. 
We found no evidence for such large   phase delay effect in modern
software (in 171 software projects shepherded
by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
Whatever the reason, the main point  is that phase delay is a widely
held belief, despite little recent evidence to support it.
\item
Devanbu et al.  examined responses from 564 Microsoft software developers from around
the world, They found that  ``(a)~programmers do indeed have very
strong beliefs on certain topics; (b)~their beliefs are primarily formed
based on personal experience, rather than on findings in empirical
research; (c)~beliefs can vary with each project, but do not necessarily
correspond with actual evidence in that project''~\cite{prem16}.
\end{enumerate}
Devanbu et al. further  comment that ``programmers give personal experience
as the strongest influence in forming their opinions''. This is a troubling
result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
careers.

If the above remarks hold true for bad smells, then we would expect
to see much disagreement on what bad smells are important and relevant
to  a particular project. This is indeed the case.
The first column of \fig{smells} 
lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{Fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
The other
columns show other work that comments on which bad smells matter most.
The columns marked as Lanza'06, Yamashita'13, and Atwood'06 come from
peer reviewed literature. The column maked SonarQube is an open source
code assessment tool that includes dectectors for six of the bad smells
in column one. 
The {\em developer survey}   (in the right-hand-side column),
   this shows the results of a hour-long whiteboard sessions with a group of a
 dozen developers from one Washington
    D.C. web-tools development organization. In that study, participants
    worked in a round robin manner to rank the bad smells they thought were
    important (and any disagreements were discussed with the whole group).
     Amongst those group, there was there was  some
    consensus on  the priority of fixing which  bad smells   
    (see the annotations VH=very high,
    H=high, L=low, VL=very low, and ``?''= no consensus).

A  blank cell in \fig{smells}
indicates where   other work has chosen to ignore
one the   bad smells in column one. 
Note that most of the cells are blank.
While the Atwood'06~\cite{Atwood06} study  comments on   
all the   bad smells,
all the  other  studies ignore the majority of the Fowler bad smells.
For exampel SonarQube has no detectors for many of the column one bad smells.
Also, nearly half the Yamashita list of bad smells
    does not appear in the Fowler and Kerievsky list
    (the eight numbers
    in the  Yamashita'13 column show the rankings for the bad smells 
    that overlap with Fowler and Kerievsky).

Two of the studies in \fig{smells} offers some comments on the relative importance
of the different bad smells the   Yamashita'13 study and the developer study:
    \begin{itemize}
        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
        and {\em long method}.
        \item
 Note that this agreement also means that the
  Yamashita'13 study and the {\em developer survey}   concluded
        that most of the  listed code smells are {\em not} high priority issues
        requiring code reorganization. 
 \end{itemize} 
In summary, just because a developer strongly believes something
about bad smells, does
not necessary mean that that belief is useful for the current project.
Human developers can be clever, but their thinking can also be distorted
by congitive biases.
Hence it is important to double check their beliefs in, for example,
bad smells. Tools like XTREE perform that check, using models learned
from the historical record.

\input{tex/jur_char}

### 2016-03-08 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|l|c|c|c|c}
Mostly from Fowler'99& & & & & Developer\\
 (with a few additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey\\\hline
  Alt. Classes with Diff. Interfaces & & & & + & \\
  Combinatorial Explosion~\cite{Kerievsky2004} & & & & + & \\
  Comments & & & 11 & + & VL\\
  Conditional Complexity~\cite{Kerievsky2004} & & & 14 & + & ?\\
  Data Class & Data Class & & & + &\\
  Data Clumps &  &  & & + &\\
  Divergent Change & & & & + & \\
  Duplicated Code & Significant Duplication\textsuperscript{*} & + & 1 & + & VH\\
  Feature Envy & Feature Envy & & 8 & + &\\

  Inappropriate Intimacy & & - & & + & L\\
  Indecent Exposure~\cite{Kerievsky2004} & & & & + & ?\\
  Incomplete Library Class & & & & + &\\
  Large Class & God Class/Brain Class & ++ & 4 & + & VH\\
  Lazy Class/Freeloader & & - & 7 & + &\\
  Long Method & Brain Method & ++ & 2 & + & VH\\
  Long Parameter List &  & + & 9 & + & L \\

  Message Chains & & & & + & H\\
  Middle Man & &  & & + &\\
  Oddball Solution~\cite{Kerievsky2004} & & & & + & \\
  Parallel Inheritance Hierarchies & & & & + &\\
  Primitive Obsession &  & & & + &\\

  Refused Bequest & Refused Parent Bequest & - & & + & \\ 
  Shotgun Surgery & Shotgun Surgery & & & + & \\
  Solution Sprawl~\cite{Kerievsky2004} & & & & + &\\
  Speculative Generality & & & & + & L\\
  Switch Statements &  & - & & & L\\
  Temporary Field & & - & & + & ?\\
  \end{tabular}
%   \multicolumn{5}{l}{} \\
%   \multicolumn{5}{l}{\textsuperscript{\textdagger} + Exact rule, - Possibly related rule} \\
%     \multicolumn{5}{l}{\textsuperscript{\textdaggerdbl} + is included in list} \\
%     \multicolumn{5}{l}{\textsuperscript{\textexclamdown} VH: Very High, H: High, L: Low, VL: Very Low, ?: Still investigating} \\
%   \multicolumn{5}{l}{\textsuperscript{*} Computation is based on more than simple metrics.} \\

\caption{List of bad code smells from different sources. Empty cells
denote some bad smell proposed by Fowler'99 that was not found relevant
in subsequent studies. Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}

>>>>

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|c|c|c|c|c}
Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
 (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
  Comments & & & 11 & \checkmark & VL\\
  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
  Data Class & \checkmark & & & \checkmark &\\
  Data Clumps &  &  & & \checkmark &\\
  Divergent Change & & & & \checkmark & \\
  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
  Feature Envy & \checkmark & & 8 & \checkmark &\\

  Inappropriate Intimacy & & \checkmark & & \checkmark & L\\
  Indecent Exposure~\cite{Kerievsky2005} & & & & \checkmark & ?\\
  Incomplete Library Class & & & & \checkmark &\\
  Large Class & \checkmark & \checkmark & 4 & \checkmark & VH\\
  Lazy Class/Freeloader & & \checkmark & 7 & \checkmark &\\
  Long Method & \checkmark& \checkmark & 2 & \checkmark & VH\\
  Long Parameter List &  & \checkmark & 9 & \checkmark & L \\

  Message Chains & & & & \checkmark & H\\
  Middle Man & &  & & \checkmark &\\
  Oddball Solution~\cite{Kerievsky2005} & & & & \checkmark & \\
  Parallel Inheritance Hierarchies & & & & \checkmark &\\
  Primitive Obsession &  & & & \checkmark &\\

  Refused Bequest & \checkmark & \checkmark & & \checkmark & \\ 
  Shotgun Surgery & \checkmark& & & \checkmark & \\
  Solution Sprawl~\cite{Kerievsky2005} & & & & \checkmark &\\
  Speculative Generality & & & & \checkmark & L\\
  Switch Statements &  & \checkmark & & & L\\
  Temporary Field & & \checkmark & & \checkmark & ?\\
  \end{tabular}
%   \multicolumn{5}{l}{} \\
%   \multicolumn{5}{l}{\textsuperscript{\textdagger} + Exact rule, - Possibly related rule} \\
%     \multicolumn{5}{l}{\textsuperscript{\textdaggerdbl} + is included in list} \\
%     \multicolumn{5}{l}{\textsuperscript{\textexclamdown} VH: Very High, H: High, L: Low, VL: Very Low, ?: Still investigating} \\
%   \multicolumn{5}{l}{\textsuperscript{*} Computation is based on more than simple metrics.} \\

\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
``?'' indicates lack of consensus on some bad smell. Empty cells
denote some bad smell listed in column one that was not found relevant
in subsequent studies.
Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}
